Weighted Model Counting with Conditional Weights

### 2020-08-02 paper/paper.tex
Two-column mode

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\usepackage{fullpage}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage[capitalise]{cleveref}
\usepackage{bm}
\usepackage{booktabs}
\usepackage[table]{xcolor}
\usepackage{tikz}
\usepackage[backgroundcolor=lightgray]{todonotes}
\usepackage{complexity}
\usepackage{subcaption}
\usepackage{multirow}
\usepackage[binary-units]{siunitx}
\usepackage[inline]{enumitem}

\usetikzlibrary{cd}
\usetikzlibrary{bayesnet}
\usetikzlibrary{calc}

\newtheorem{theorem}{Theorem}
\newtheorem{observation}{Observation}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\DeclareMathOperator{\im}{im}

\title{Weighted Model Counting with Conditional Weights}
\author{Paulius Dilkas}

@@

@@

\begin{itemize}
\item The main narrative
  \begin{enumerate}
  \item Putting weights on literals is restrictive.
  \item To overcome this, it is common practice to add more literals.
  \item This is indeed always possible but much slower.
  \item We show how we can define conditional weights on literals, allowing us
    to encode any measure.
  \item We demonstrate a specific example of this by presenting a new way to
    encode Bayesian networks into instances of WMC and adapting a WMC algorithm
    (ADDMC) to run on the new format.
  \item We show that this results in significantly faster inference.
  \item We show that our encoding results in asymptotically fewer literals and
    fewer ADDs.
  \end{enumerate}
\item Introduce
  \begin{itemize}
  \item WMC \cite{DBLP:conf/aaai/SangBK05}
  \item ADDMC \cite{DBLP:conf/aaai/DudekPV20} and ADDs
    \cite{DBLP:journals/fmsd/BaharFGHMPS97}
  \item A Bayesian network is a directed acyclic graph with random variables as
    vertices that defines a probability distribution over them. The full
    probability distribution can be expressed as the product of the conditional
    probability distributions of all random variables. For discrete Bayesian
    networks (and we only consider discrete networks in this paper), each such
    distribution can be represented by a table known as a conditional
    probability table (CPT).
  \end{itemize}
\item[F] What are the main claims, what are the main takeaways, intuitive [???]
  of theorems to follow.
\item Our work is, in spirit, similar to... using ROBDDs for Inference in
  Bayesian Networks with Troubleshooting as an Example
  \cite{DBLP:conf/uai/NielsenWJK00}: this is an interesting approach where
  deterministic parts of a BN are expressed as Boolean functions, i.e.,
  extracting the logical from the probabilistic. Maybe I can make a case that my
  encoding (the textual version) kind of does the same, even though the
  algorithm doesn't.
\end{itemize}

@@

\paragraph{Using WMC to perform inference on Bayesian networks.} Hitherto, four
techniques have been proposed for encoding Bayesian networks into instances of
WMC. We will identify them based on the initials of authors as well as
publications years: \texttt{d02} \cite{DBLP:conf/kr/Darwiche02}, \texttt{sbk05}
\cite{DBLP:conf/aaai/SangBK05}, \texttt{cd05} \cite{DBLP:conf/ijcai/ChaviraD05},
and \texttt{cd06} \cite{DBLP:conf/sat/ChaviraD06}. Below we summarise the
observed performance differences among them. Sang et al.
\cite{DBLP:conf/aaai/SangBK05} claim that \texttt{sbk05} is a smaller encoding
than \texttt{d02} with respect to both the number of clauses and the number of
variables but provide no experimental comparison. Chavira and Darwiche
\cite{DBLP:conf/ijcai/ChaviraD05} compare \texttt{cd05} with \texttt{d02} by
measuring the time it takes to compile either encoding into an arithmetic
circuit (but do not measure inference time). The more recent encoding
\texttt{cd05} always compiles faster and results in a smaller arithmetic circuit
(as measured by the number of edges). In their subsequent paper, the same
authors perform two sets of experiments (that are relevant to this summary)
\cite{DBLP:conf/sat/ChaviraD06}. First, they compile \texttt{cd05} and
\texttt{cd06} encodings into d-DNNF (i.e., deterministic decomposable negation
normal form \cite{DBLP:journals/jancl/Darwiche01}), measuring both compilation
time and numbers of edges in the d-DNNF diagram. The results are mostly in
favour of \texttt{cd06}. Second, they compare the inference time of
\texttt{sbk05} run with Cachet \cite{DBLP:conf/sat/SangBBKP04} with the compile
times of \texttt{cd05} and \texttt{cd06}, but only on five (types of) instances.
In these experiments, \texttt{cd06} is always faster than \texttt{cd05}, while
the comparison with \texttt{sbk05} is mixed. Sometimes \texttt{cd06} is orders
of magnitude faster than \texttt{sbk05}, sometimes slightly slower. The
performance difference between \texttt{sbk05} and \texttt{cd05} is even harder
to judge: \texttt{sbk05} is better on three out of five instances and worse on
the remaining two. Based on this description, one would expect \texttt{cd06} to
be faster than both \texttt{cd05} and \texttt{sbk05}, both of which should be
faster than \texttt{d02}. The experiments in \cref{sec:experiments}, however,
strongly disagree with this prediction, showing that the quality of an encoding
depends strongly on the underlying search algorithm or compilation technique.

@@

\begin{table}
  \centering
  \caption{A comparison of Boolean-algebraic (BA) and set-theoretic (ST)
    concepts for $2^U$ for any set $U$}
  \label{tbl:notation}
  \begin{tabular}{lccl}
    \toprule
    BA name & BA symbol & ST symbol & ST name \\
    \midrule
    bottom & $\bot$ & $\emptyset$ & empty set \\
    top & $\top$ & $U$ & \\
    meet, and & $\land$ & $\cap$ & intersection \\
    join, or & $\lor$ & $\cup$ & union \\
    complement, not & $\neg$ & $^c$ & complement \\
            & $\le$ & $\subseteq$ & subset relation, set inclusion \\
    atom & & & singleton, unit set \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}
  \caption{Notation for a logic with two atoms. The elements in both columns are
    listed in the same order.}
  \label{tbl:notation_example}
  \centering
  \begin{tabular}{lcc}
    \toprule
    Name in logic & Boolean-algebraic notation & Set-theoretic notation \\
    \midrule
    Atoms (elements of $U$) & $a, b$ & $a, b$ \\
    \rowcolor{gray!10} Models (elements of $2^U$) & $\neg a \land \neg b, a \land \neg b, \neg a \land b, a \land b$ & $\emptyset, \{a\}, \{b\}, \{a, b\}$ \\
    & $\top$ & $\{ \emptyset, \{a\}, \{b\}, \{a, b\} \}$ \\
    & $\neg a \lor \neg b, a \to b$ & $\{ \emptyset, \{a\}, \{b\} \}, \{ \emptyset, \{a\}, \{a, b\} \}$ \\
    & $b \to a, a \lor b$ & $\{ \emptyset, \{b\}, \{a, b\} \}, \{ \{a\}, \{b\}, \{a, b\} \}$ \\
    & $\neg b, \neg a, a \leftrightarrow b$ & $\{\emptyset, \{a\}\}, \{\emptyset, \{b\}\}, \{\emptyset, \{a, b\}\}$ \\
    & $(a \land \neg b) \lor (b \land \neg a), a, b$ & $\{\{a\}, \{b\}\}, \{\{a\}, \{a, b\}\}, \{\{b\}, \{a, b\}\}$ \\
    & $\neg a \land \neg b, a \land \neg b, \neg a \land b, a \land b$ & $\{\emptyset\}, \{\{a\}\}, \{\{b\}\}, \{\{a, b\}\}$ \\
    \multirow{-7}{*}{Formulas (elements of $2^{2^U}$)} & $\bot$ & $\emptyset$ \\
    \bottomrule
  \end{tabular}
\end{table}

@@

\begin{lemma} \label{prop:measure}
  The function $\mu_\nu$, as defined by \cref{eq:induced_measure}, is a measure.
\end{lemma}
\begin{proof}
  Note that $\mu_\nu(\bot) = 0$ since there are no atoms below $\bot$. Let $a, b
  \in 2^{2^{U}}$ be such that $a \land b = \bot$. By elementary properties of
  Boolean algebras, all atoms below $a \lor b$ are either below $a$ or below
  $b$. Moreover, none of them can be below both $a$ and $b$ because then they
  would have to be below $a \land b = \bot$. Thus
  \[
    \mu_\nu(a \lor b) = \sum_{\{u\} \le a \lor b} \nu(u) = \sum_{\{u\} \le a}
    \nu(u) + \sum_{\{u\} \le b} \nu(u) = \mu_\nu(a) + \mu_\nu(b)
  \]
  as required.
\end{proof}

@@

\paragraph{Relation to the classical (logic-based) view of WMC.} Let
$\mathcal{L}$ be a propositional logic with two atoms $a$ and $b$ as in
\cref{sec:prelims} and $w\colon \{ a, b, \neg a, \neg b \} \to \mathbb{R}_{\ge
  0}$ a \emph{weight function} defined as $w(a) = 0.3$, $w(\neg a) = 0.7$, $w(b)
= 0.2$, $w(\neg b) = 0.8$. Furthermore, let $\Delta$ be a theory in
$\mathcal{L}$ with a sole axiom $a$. Then $\Delta$ has two models: $\{ a, b \}$
and $\{ a, \neg b \}$ and its WMC \cite{DBLP:journals/ai/ChaviraD08} is
\begin{equation} \label{eq:wmc_example}
  \mathrm{WMC}(\Delta) = \sum_{\omega \models \Delta} \prod_{\omega \models l} w(l) = w(a)w(b) + w(a)w(\neg b) = 0.3.
\end{equation}
Alternatively, we can define $\nu_a\colon 2^{\{a\}} \to \mathbb{R}_{\ge 0}$ as
$\nu_a(\{ a \}) = 0.3$, $\nu_a(\emptyset) = 0.7$ and $\nu_b\colon 2^{\{b\}} \to
\mathbb{R}_{\ge 0}$ as $\nu_b(\{ b \}) = 0.2$, $\nu_b(\emptyset) = 0.8$. Let
$\mu$ be the measure on $2^{2^U}$ induced by $\nu = \nu_a \cdot \nu_b$. Then,
equivalently to \cref{eq:wmc_example}, we can write
\[
  \mu(a) = \nu(\{ a, b \}) + \nu(\{ a \}) = \nu_a(\{a\})\nu_b(\{b\}) +
  \nu_a(\{a\})\nu_b(\emptyset) = 0.3.
\]

@@

\begin{figure}
  \centering
  \begin{subfigure}{0.2\textwidth}
    \centering
    \begin{tikzpicture}[edge from parent/.style={draw,-latex}]
      \node[draw,circle] {$W$}
      child {node[draw,circle] {$F$}}
      child {node[draw,circle] {$T$}};
    \end{tikzpicture}
  \end{subfigure}%
  \begin{subfigure}{0.8\textwidth}
    \centering
    \begin{tabular}[t]{cc}
      \toprule
      $w$ & $\Pr(W = w)$ \\
      \midrule
      1 & 0.5 \\
      0 & 0.5 \\
      \bottomrule
    \end{tabular}
    \begin{tabular}[t]{ccc}
      \toprule
      $w$ & $f$ & $\Pr(F = f \mid W = w)$ \\
      \midrule
      1 & 1 & 0.6 \\
      1 & 0 & 0.4 \\
      0 & 1 & 0.1 \\
      0 & 0 & 0.9 \\
      \bottomrule
    \end{tabular}
    \begin{tabular}[t]{ccc}
      \toprule
      $w$ & $t$ & $\Pr(T = t \mid W = w)$ \\
      \midrule
      1 & $l$ & 0.2 \\
      1 & $m$ & 0.4 \\
      1 & $h$ & 0.4 \\
      0 & $l$ & 0.6 \\
      0 & $m$ & 0.3 \\
      0 & $h$ & 0.1 \\
      \bottomrule
    \end{tabular}
  \end{subfigure}
  \caption{An example Bayesian network with its CPTs}
  \label{fig:example_bn}
\end{figure}

@@

\begin{algorithm}
  \caption{Encoding a Bayesian network as a function $2^U \to \mathbb{R}_{\ge
      0}$}
  \label{alg:encoding}
  \KwData{a Bayesian network with vertices $\mathcal{V}$ and probability
    distribution $\Pr$}
  \KwResult{a function $\phi\colon 2^U \to \mathbb{R}_{\ge 0}$}
  $\phi \gets 1$\;
  \For{$X \in \mathcal{V}$}{
    \textit{let} $\mathrm{pa}(X) = \{ Y_1, \dots, Y_n \}$\;
    $\mathrm{CPT}_X \gets 0$\;
    \eIf{$|\im X| = 2$}{
      \For{$(y_1, \dots, y_n) \in \prod_{i = 1}^n \im Y_i$}{
        $p_1 \gets \Pr(X = 1 \mid Y_1 = y_1, \dots, Y_n = y_n)$\;
        $p_0 \gets \Pr(X \ne 1 \mid Y_1 = y_1, \dots, Y_n = y_n)$\;
        $\mathrm{CPT}_X \gets \mathrm{CPT}_X + p_1[\lambda_{X=1}] \cdot
        \prod_{i=1}^n [\lambda_{Y_i=y_i}] + p_0 \overline{[\lambda_{X=1}]} \cdot
        \prod_{i=1}^n [\lambda_{Y_i=y_i}]$\;
      }
    }{
      \textit{let} $\im X = \{ x_1, \dots, x_m \}$\;
      \For{$x \in \im X$ {\rm \textbf{and}} $(y_1, \dots, y_n) \in \prod_{i =
          1}^n \im Y_i$}{
        $p_x \gets \Pr(X = x \mid Y_1 = y_1, \dots, Y_n = y_n)$\;
        $\mathrm{CPT}_X \gets \mathrm{CPT}_X + p_x[\lambda_{X=x}] \cdot
        \prod_{i=1}^n [\lambda_{Y_i=y_i}] + \overline{[\lambda_{X=1}]} \cdot
        \prod_{i=1}^n [\lambda_{Y_i=y_i}]$\;
      }
      $\mathrm{CPT}_X \gets \mathrm{CPT}_X \cdot \left( \sum_{i=1}^m [\lambda_{X
          = x_i}] \right) \cdot \prod_{i=1}^m \prod_{j=i+1}^m
      (\overline{[\lambda_{X = x_i}]} + \overline{[\lambda_{X = x_j}]})$\;
    }
    $\phi \gets \phi \cdot \mathrm{CPT}_X$\;
  }
  \Return{$\phi$}\;
\end{algorithm}

>>>>

\def\year{2020}\relax
%File: formatting-instruction.tex
\documentclass[letterpaper]{article} % DO NOT CHANGE THIS
\usepackage{aaai20}  % DO NOT CHANGE THIS
\usepackage{times}  % DO NOT CHANGE THIS
\usepackage{helvet} % DO NOT CHANGE THIS
\usepackage{courier}  % DO NOT CHANGE THIS
\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
\usepackage{graphicx} % DO NOT CHANGE THIS
\urlstyle{rm} % DO NOT CHANGE THIS
\def\UrlFont{\rm}  % DO NOT CHANGE THIS
\usepackage{graphicx}  % DO NOT CHANGE THIS
\frenchspacing  % DO NOT CHANGE THIS
\setlength{\pdfpagewidth}{8.5in}  % DO NOT CHANGE THIS
\setlength{\pdfpageheight}{11in}  % DO NOT CHANGE THIS
%\nocopyright
%PDF Info Is REQUIRED.
% For /Author, add all authors within the parentheses, separated by commas. No accents or commands.
% For /Title, add Title in Mixed Case. No accents or commands. Retain the parentheses.
 \pdfinfo{
/Title (AAAI Press Formatting Instructions for Authors Using LaTeX -- A Guide)
/Author (AAAI Press Staff, Pater Patel Schneider, Sunil Issar, J. Scott Penberthy, George Ferguson, Hans Guesgen)
} %Leave this	
% /Title ()
% Put your actual complete title (no codes, scripts, shortcuts, or LaTeX commands) within the parentheses in mixed case
% Leave the space between \Title and the beginning parenthesis alone
% /Author ()
% Put your actual complete list of authors (no codes, scripts, shortcuts, or LaTeX commands) within the parentheses in mixed case. 
% Each author should be only by a comma. If the name contains accents, remove them. If there are any LaTeX commands, 
% remove them.

% DISALLOWED PACKAGES
% \usepackage{authblk} -- This package is specifically forbidden
% \usepackage{balance} -- This package is specifically forbidden
% \usepackage{caption} -- This package is specifically forbidden
% \usepackage{color (if used in text)
% \usepackage{CJK} -- This package is specifically forbidden
% \usepackage{float} -- This package is specifically forbidden
% \usepackage{flushend} -- This package is specifically forbidden
% \usepackage{fontenc} -- This package is specifically forbidden
% \usepackage{fullpage} -- This package is specifically forbidden
% \usepackage{geometry} -- This package is specifically forbidden
% \usepackage{grffile} -- This package is specifically forbidden
% \usepackage{hyperref} -- This package is specifically forbidden
% \usepackage{navigator} -- This package is specifically forbidden
% (or any other package that embeds links such as navigator or hyperref)
% \indentfirst} -- This package is specifically forbidden
% \layout} -- This package is specifically forbidden
% \multicol} -- This package is specifically forbidden
% \nameref} -- This package is specifically forbidden
% \natbib} -- This package is specifically forbidden -- use the following workaround:
% \usepackage{savetrees} -- This package is specifically forbidden
% \usepackage{setspace} -- This package is specifically forbidden
% \usepackage{stfloats} -- This package is specifically forbidden
% \usepackage{tabu} -- This package is specifically forbidden
% \usepackage{titlesec} -- This package is specifically forbidden
% \usepackage{tocbibind} -- This package is specifically forbidden
% \usepackage{ulem} -- This package is specifically forbidden
% \usepackage{wrapfig} -- This package is specifically forbidden
% DISALLOWED COMMANDS
% \nocopyright -- Your paper will not be published if you use this command
% \addtolength -- This command may not be used
% \balance -- This command may not be used
% \baselinestretch -- Your paper will not be published if you use this command
% \clearpage -- No page breaks of any kind may be used for the final version of your paper
% \columnsep -- This command may not be used
% \newpage -- No page breaks of any kind may be used for the final version of your paper
% \pagebreak -- No page breaks of any kind may be used for the final version of your paperr
% \pagestyle -- This command may not be used
% \tiny -- This is not an acceptable font size.
% \vspace{- -- No negative value may be used in proximity of a caption, figure, table, section, subsection, subsubsection, or reference
% \vskip{- -- No negative value may be used to alter spacing above or below a caption, figure, table, section, subsection, subsubsection, or reference

\setcounter{secnumdepth}{2} %May be changed to 1 or 2 if section numbers are desired.

\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage[capitalise]{cleveref}
\usepackage{bm}
\usepackage{booktabs}
\usepackage[table]{xcolor}
\usepackage{tikz}
\usepackage[backgroundcolor=lightgray]{todonotes}
\usepackage{complexity}
\usepackage{subcaption}
\usepackage{multirow}
\usepackage[binary-units]{siunitx}
\usepackage[inline]{enumitem}
\usetikzlibrary{cd}
\usetikzlibrary{bayesnet}
\usetikzlibrary{calc}
\DeclareMathOperator{\im}{im}
\newtheorem{theorem}{Theorem}
\newtheorem{observation}{Observation}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

% The file aaai20.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\setlength\titlebox{2.5in} % If your paper contains an overfull \vbox too high warning at the beginning of the document, use this
% command to correct it. You may not alter the value below 2.5 in
\title{Weighted Model Counting with Conditional Weights}
%Your title must be in mixed case, not sentence case. 
% That means all verbs (including short verbs like be, is, using,and go), 
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
\author{Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{Primarily Mike Hamilton of the Live Oak Press, LLC, with help from the AAAI Publications Committee}\\ \Large \textbf{AAAI Style Contributions by
Pater Patel Schneider,} \\ \Large \textbf{Sunil Issar, J. Scott Penberthy, George Ferguson, Hans Guesgen}\\ % All authors must be in the same font size and format. Use \Large and \textbf to achieve this result when breaking a line
\textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\ %If you have multiple authors and multiple affiliations
% use superscripts in text and roman font to identify them. For example, Sunil Issar,\textsuperscript{\rm 2} J. Scott Penberthy\textsuperscript{\rm 3} George Ferguson,\textsuperscript{\rm 4} Hans Guesgen\textsuperscript{\rm 5}. Note that the comma should be placed BEFORE the superscript for optimum readability
2275 East Bayshore Road, Suite 160\\
Palo Alto, California 94303\\
publications20@aaai.org % email address must be in roman text type, not monospace or sans serif
}

@@

\begin{abstract}
AAAI creates proceedings, working notes, and technical reports directly from
electronic source furnished by the authors. To ensure that all papers in the
publication have a uniform appearance, authors must adhere to the following
instructions.
\end{abstract}

@@

\noindent Foo
\begin{itemize}
\item The main narrative
  \begin{enumerate}
  \item Putting weights on literals is restrictive.
  \item To overcome this, it is common practice to add more literals.
  \item This is indeed always possible but much slower.
  \item We show how we can define conditional weights on literals, allowing us
    to encode any measure.
  \item We demonstrate a specific example of this by presenting a new way to
    encode Bayesian networks into instances of WMC and adapting a WMC algorithm
    (ADDMC) to run on the new format.
  \item We show that this results in significantly faster inference.
  \item We show that our encoding results in asymptotically fewer literals and
    fewer ADDs.
  \end{enumerate}
\item Introduce
  \begin{itemize}
  \item WMC \cite{DBLP:conf/aaai/SangBK05}
  \item ADDMC \cite{DBLP:conf/aaai/DudekPV20} and ADDs
    \cite{DBLP:journals/fmsd/BaharFGHMPS97}
  \item A Bayesian network is a directed acyclic graph with random variables as
    vertices that defines a probability distribution over them. The full
    probability distribution can be expressed as the product of the conditional
    probability distributions of all random variables. For discrete Bayesian
    networks (and we only consider discrete networks in this paper), each such
    distribution can be represented by a table known as a conditional
    probability table (CPT).
  \end{itemize}
\item[F] What are the main claims, what are the main takeaways, intuitive [???]
  of theorems to follow.
\item Our work is, in spirit, similar to... using ROBDDs for Inference in
  Bayesian Networks with Troubleshooting as an Example
  \cite{DBLP:conf/uai/NielsenWJK00}: this is an interesting approach where
  deterministic parts of a BN are expressed as Boolean functions, i.e.,
  extracting the logical from the probabilistic. Maybe I can make a case that my
  encoding (the textual version) kind of does the same, even though the
  algorithm doesn't.
\item Replace things like `Darwiche (Darwiche, 2005)' with `Darwiche (2005)' if
  possible (without natbib).
\end{itemize}

@@

\paragraph{WMC for Bayesian networks.} Hitherto, four techniques have been
proposed for encoding Bayesian networks into instances of WMC. We will identify
them based on the initials of authors as well as publications years:
\texttt{d02} \cite{DBLP:conf/kr/Darwiche02}, \texttt{sbk05}
\cite{DBLP:conf/aaai/SangBK05}, \texttt{cd05} \cite{DBLP:conf/ijcai/ChaviraD05},
and \texttt{cd06} \cite{DBLP:conf/sat/ChaviraD06}. Below we summarise the
observed performance differences among them. Sang et al.
\cite{DBLP:conf/aaai/SangBK05} claim that \texttt{sbk05} is a smaller encoding
than \texttt{d02} with respect to both the number of clauses and the number of
variables but provide no experimental comparison. Chavira and Darwiche
\cite{DBLP:conf/ijcai/ChaviraD05} compare \texttt{cd05} with \texttt{d02} by
measuring the time it takes to compile either encoding into an arithmetic
circuit (but do not measure inference time). The more recent encoding
\texttt{cd05} always compiles faster and results in a smaller arithmetic circuit
(as measured by the number of edges). In their subsequent paper, the same
authors perform two sets of experiments (that are relevant to this summary)
\cite{DBLP:conf/sat/ChaviraD06}. First, they compile \texttt{cd05} and
\texttt{cd06} encodings into d-DNNF (i.e., deterministic decomposable negation
normal form \cite{DBLP:journals/jancl/Darwiche01}), measuring both compilation
time and numbers of edges in the d-DNNF diagram. The results are mostly in
favour of \texttt{cd06}. Second, they compare the inference time of
\texttt{sbk05} run with Cachet \cite{DBLP:conf/sat/SangBBKP04} with the compile
times of \texttt{cd05} and \texttt{cd06}, but only on five (types of) instances.
In these experiments, \texttt{cd06} is always faster than \texttt{cd05}, while
the comparison with \texttt{sbk05} is mixed. Sometimes \texttt{cd06} is orders
of magnitude faster than \texttt{sbk05}, sometimes slightly slower. The
performance difference between \texttt{sbk05} and \texttt{cd05} is even harder
to judge: \texttt{sbk05} is better on three out of five instances and worse on
the remaining two. Based on this description, one would expect \texttt{cd06} to
be faster than both \texttt{cd05} and \texttt{sbk05}, both of which should be
faster than \texttt{d02}. The experiments in \cref{sec:experiments}, however,
strongly disagree with this prediction, showing that the quality of an encoding
depends strongly on the underlying search algorithm or compilation technique.

@@

\todo[inline]{Describe BA-PS duality without the table.}

% \begin{table}
%   \centering
%   \caption{A comparison of Boolean-algebraic (BA) and set-theoretic (ST)
%     concepts for $2^U$ for any set $U$}
%   \label{tbl:notation}
%   \begin{tabular}{lccl}
%     \toprule
%     BA name & BA symbol & ST symbol & ST name \\
%     \midrule
%     bottom & $\bot$ & $\emptyset$ & empty set \\
%     top & $\top$ & $U$ & \\
%     meet, and & $\land$ & $\cap$ & intersection \\
%     join, or & $\lor$ & $\cup$ & union \\
%     complement, not & $\neg$ & $^c$ & complement \\
%             & $\le$ & $\subseteq$ & subset relation, set inclusion \\
%     atom & & & singleton, unit set \\
%     \bottomrule
%   \end{tabular}
% \end{table}

\begin{table*}
  \caption{Notation for a logic with two atoms. The elements in both columns are
    listed in the same order.}
  \label{tbl:notation_example}
  \centering
  \begin{tabular}{lcc}
    \toprule
    Name in logic & Boolean-algebraic notation & Set-theoretic notation \\
    \midrule
    Atoms (elements of $U$) & $a, b$ & $a, b$ \\
    \rowcolor{gray!10} Models (elements of $2^U$) & $\neg a \land \neg b, a \land \neg b, \neg a \land b, a \land b$ & $\emptyset, \{a\}, \{b\}, \{a, b\}$ \\
    & $\top$ & $\{ \emptyset, \{a\}, \{b\}, \{a, b\} \}$ \\
    & $\neg a \lor \neg b, a \to b$ & $\{ \emptyset, \{a\}, \{b\} \}, \{ \emptyset, \{a\}, \{a, b\} \}$ \\
    & $b \to a, a \lor b$ & $\{ \emptyset, \{b\}, \{a, b\} \}, \{ \{a\}, \{b\}, \{a, b\} \}$ \\
    & $\neg b, \neg a, a \leftrightarrow b$ & $\{\emptyset, \{a\}\}, \{\emptyset, \{b\}\}, \{\emptyset, \{a, b\}\}$ \\
    & $(a \land \neg b) \lor (b \land \neg a), a, b$ & $\{\{a\}, \{b\}\}, \{\{a\}, \{a, b\}\}, \{\{b\}, \{a, b\}\}$ \\
    & $\neg a \land \neg b, a \land \neg b, \neg a \land b, a \land b$ & $\{\emptyset\}, \{\{a\}\}, \{\{b\}\}, \{\{a, b\}\}$ \\
    \multirow{-7}{*}{Formulas (elements of $2^{2^U}$)} & $\bot$ & $\emptyset$ \\
    \bottomrule
  \end{tabular}
\end{table*}

@@

\begin{lemma} \label{prop:measure}
  The function $\mu_\nu$, (from \cref{eq:induced_measure}), is a measure.
\end{lemma}
% \begin{proof}
%   Note that $\mu_\nu(\bot) = 0$ since there are no atoms below $\bot$. Let $a, b
%   \in 2^{2^{U}}$ be such that $a \land b = \bot$. By elementary properties of
%   Boolean algebras, all atoms below $a \lor b$ are either below $a$ or below
%   $b$. Moreover, none of them can be below both $a$ and $b$ because then they
%   would have to be below $a \land b = \bot$. Thus
%   \[
%     \mu_\nu(a \lor b) = \sum_{\{u\} \le a \lor b} \nu(u) = \sum_{\{u\} \le a}
%     \nu(u) + \sum_{\{u\} \le b} \nu(u) = \mu_\nu(a) + \mu_\nu(b)
%   \]
%   as required.
% \end{proof}

@@

\paragraph{Relation to the classical (logic-based) view of WMC.} Let
$\mathcal{L}$ be a propositional logic with two atoms $a$ and $b$ as in
\cref{sec:prelims} and $w\colon \{ a, b, \neg a, \neg b \} \to \mathbb{R}_{\ge
  0}$ a \emph{weight function} defined as $w(a) = 0.3$, $w(\neg a) = 0.7$, $w(b)
= 0.2$, $w(\neg b) = 0.8$. Furthermore, let $\Delta$ be a theory in
$\mathcal{L}$ with a sole axiom $a$. Then $\Delta$ has two models: $\{ a, b \}$
and $\{ a, \neg b \}$ and its WMC \cite{DBLP:journals/ai/ChaviraD08} is
\begin{equation} \label{eq:wmc_example}
  \begin{split}
    \mathrm{WMC}(\Delta) &= \sum_{\omega \models \Delta} \prod_{\omega \models l} w(l) \\
    &= w(a)w(b) + w(a)w(\neg b) = 0.3.
  \end{split}
\end{equation}
Alternatively, we can define $\nu_a\colon 2^{\{a\}} \to \mathbb{R}_{\ge 0}$ as
$\nu_a(\{ a \}) = 0.3$, $\nu_a(\emptyset) = 0.7$ and $\nu_b\colon 2^{\{b\}} \to
\mathbb{R}_{\ge 0}$ as $\nu_b(\{ b \}) = 0.2$, $\nu_b(\emptyset) = 0.8$. Let
$\mu$ be the measure on $2^{2^U}$ induced by $\nu = \nu_a \cdot \nu_b$. Then,
equivalently to \cref{eq:wmc_example}, we can write
\begin{align*}
  \mu(a) &= \nu(\{ a, b \}) + \nu(\{ a \}) \\
         &= \nu_a(\{a\})\nu_b(\{b\}) + \nu_a(\{a\})\nu_b(\emptyset) = 0.3.
\end{align*}

@@

\begin{figure*}
  \centering
  \begin{subfigure}{0.2\textwidth}
    \centering
    \begin{tikzpicture}[edge from parent/.style={draw,-latex}]
      \node[draw,circle] {$W$}
      child {node[draw,circle] {$F$}}
      child {node[draw,circle] {$T$}};
    \end{tikzpicture}
  \end{subfigure}%
  \begin{subfigure}{0.8\textwidth}
    \centering
    \begin{tabular}[t]{cc}
      \toprule
      $w$ & $\Pr(W = w)$ \\
      \midrule
      1 & 0.5 \\
      0 & 0.5 \\
      \bottomrule
    \end{tabular}
    \begin{tabular}[t]{ccc}
      \toprule
      $w$ & $f$ & $\Pr(F = f \mid W = w)$ \\
      \midrule
      1 & 1 & 0.6 \\
      1 & 0 & 0.4 \\
      0 & 1 & 0.1 \\
      0 & 0 & 0.9 \\
      \bottomrule
    \end{tabular}
    \begin{tabular}[t]{ccc}
      \toprule
      $w$ & $t$ & $\Pr(T = t \mid W = w)$ \\
      \midrule
      1 & $l$ & 0.2 \\
      1 & $m$ & 0.4 \\
      1 & $h$ & 0.4 \\
      0 & $l$ & 0.6 \\
      0 & $m$ & 0.3 \\
      0 & $h$ & 0.1 \\
      \bottomrule
    \end{tabular}
  \end{subfigure}
  \caption{An example Bayesian network with its CPTs}
  \label{fig:example_bn}
\end{figure*}

@@

\begin{algorithm}
  \caption{Encoding a Bayesian network}
  \label{alg:encoding}
  \KwData{vertices $\mathcal{V}$, probability distribution $\Pr$}
  \KwResult{a function $\phi\colon 2^U \to \mathbb{R}_{\ge 0}$}
  $\phi \gets 1$\;
  \For{$X \in \mathcal{V}$}{
    \textit{let} $\mathrm{pa}(X) = \{ Y_1, \dots, Y_n \}$\;
    $\mathrm{CPT}_X \gets 0$\;
    \eIf{$|\im X| = 2$}{
      \For{$(y_1, \dots, y_n) \in \prod_{i = 1}^n \im Y_i$}{
        $p_1 \gets \Pr(X = 1 \mid y_1, \dots, y_n)$\;
        $p_0 \gets \Pr(X \ne 1 \mid y_1, \dots, y_n)$\;
        $\mathrm{CPT}_X \gets \mathrm{CPT}_X + p_1[\lambda_{X=1}] \cdot
        \prod_{i=1}^n [\lambda_{Y_i=y_i}] + p_0 \overline{[\lambda_{X=1}]} \cdot
        \prod_{i=1}^n [\lambda_{Y_i=y_i}]$\;
      }
    }{
      \textit{let} $\im X = \{ x_1, \dots, x_m \}$\;
      \For{$x \in \im X$ {\rm \textbf{and}} $(y_1, \dots, y_n) \in \prod_{i =
          1}^n \im Y_i$}{
        $p_x \gets \Pr(X = x \mid y_1, \dots, y_n)$\;
        $\mathrm{CPT}_X \gets \mathrm{CPT}_X + p_x[\lambda_{X=x}] \cdot
        \prod_{i=1}^n [\lambda_{Y_i=y_i}] + \overline{[\lambda_{X=1}]} \cdot
        \prod_{i=1}^n [\lambda_{Y_i=y_i}]$\;
      }
      $\mathrm{CPT}_X \gets \mathrm{CPT}_X \cdot \left( \sum_{i=1}^m [\lambda_{X
          = x_i}] \right) \cdot \prod_{i=1}^m \prod_{j=i+1}^m
      (\overline{[\lambda_{X = x_i}]} + \overline{[\lambda_{X = x_j}]})$\;
    }
    $\phi \gets \phi \cdot \mathrm{CPT}_X$\;
  }
  \Return{$\phi$}\;
\end{algorithm}
